### 2025-11-25 Objects/old/hospital_general_summarystats.tex
update name cleaning and work on descriptive tables and figures

\begin{table}[ht!]

\caption{Summary Statistics of General Hospitals}
\centering
\begin{tabular}[t]{lcccc}
\toprule
Variable & Affiliated w/ General & Affiliated w/ Specialty & Unaffiliated, Part of System & Unaffiliated, Not Part of System\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Characteristics}}\\
\hspace{1em}Number of Beds & 228.79 & 689.27 & 202.39 & 117.02\\
\hspace{1em}Academic Medical Center & 0.39 & 0.76 & 0.40 & 0.29\\
\hspace{1em}In Metropolitan Area & 0.69 & 0.91 & 0.70 & 0.53\\
\hspace{1em}In System & 0.43 & 0.91 & 0.98 & 0.00\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Patients}}\\
\hspace{1em}Medicaid Patients per Bed (AHA) & 6.92 & 8.27 & 7.63 & 6.02\\
\hspace{1em}Medicare Patients per Bed (AHA) & 17.78 & 22.29 & 20.56 & 15.89\\
\hspace{1em}Number of Medicare Patients (CMS) & 3033.37 & 6281.22 & 2654.91 & 1766.75\\
\hspace{1em}Avg Medicare Patient Age & 74.43 & 74.44 & 74.13 & 74.28\\
\hspace{1em}Avg Patient Risk Score & 2.04 & 2.09 & 2.11 & 2.03\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Services}}\\
\hspace{1em}Has a Cath Lab & 0.52 & 0.88 & 0.57 & 0.36\\
\hspace{1em}Has a NICU & 0.24 & 0.61 & 0.29 & 0.14\\
\hspace{1em}Service Concentration (beds) & 0.51 & 0.48 & 0.50 & 0.56\\
\hspace{1em}Service Concentration (conditions) & 0.52 & 0.47 & 0.57 & 0.54\\
Num. Hospitals & 157.00 & 12.00 & 836.00 & 462.00\\
\bottomrule
\end{tabular}
\end{table}