### 2025-12-19 Paper/paper_2025.tex

\documentclass[12pt]{article}

\usepackage{tgtermes}
\usepackage{epsf}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue]{hyperref}
\usepackage{dcolumn}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{mwe}
\usepackage{url}
%\usepackage{harvard}
\usepackage{fancyheadings}
\usepackage{longtable}
\usepackage{authblk}
\usepackage{setspace}
%\usepackage[nomarkers]{endfloat}
\usepackage{float}
\usepackage{bbm}
%\usepackage{titling}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{import}
\usepackage[backend=biber,style=authoryear,
sorting=nyt,citestyle=authoryear]{biblatex}
\addbibresource{papercitations.bib}
%\usepackage[nomarkers,nofiglist,notablist]{endfloat}
\usepackage{subcaption}
\usepackage{caption}

\onehalfspacing
\textwidth 6.5in \oddsidemargin 0in \evensidemargin -0.6in
\textheight 8.5in \topmargin -0.2in

\newcolumntype{L}[1]{>{\raggedright\let\newline\\
		\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\
		\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\
		\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{P}[1]{>{\raggedright\tabularxbackslash}p{#1}}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}

\captionsetup{justification=centering,singlelinecheck=false}

\newcommand{\xsub}[1]{%
	\mbox{\scriptsize\begin{tabular}{@{}c@{}}#1\end{tabular}}%
}

%\renewcommand{\thetable}{\Roman{table}}

\begin{document}

	\linespread{1.2}\title{\vspace{-0.5in} Non-Merger Affiliations in Already Consolidated Health Care Markets}

	\date{\today}

	\author{\vspace{10mm}Hanna Glenn\footnote{Department of Economics, Emory University, 1602 Fishburne Drive, Atlanta, GA 30322, hanna.glenn@emory.edu.} }

	\maketitle
	%\setlength{\droptitle}{-10pt}

	\vspace{-0.2in}

	\singlespacing\maketitle

 \vspace{3mm}

    \begin{abstract}
		{\small

		} 
	\end{abstract}

	\onehalfspacing

	\newpage

    \section{Introduction}

    Market concentration has increased across industries in the last two decades (\cite{grullon2019us}), sparking renewed debate and, more recently, a shift toward stricter antitrust enforcement. Modern antitrust policy primarily focuses on mergers and acquisitions among large firms. Since the Hart–Scott–Rodino Act of 1976, the U.S. government has reviewed certain transactions that may threaten competition, defined by regulators as “a process of rivalry that incentivizes businesses to offer lower prices, improve wages and working conditions, enhance quality and resiliency, innovate, and expand choice” (\cite{DOJFTC2023MergerGuidelines}). Formal challenges remain rare: since 2015, roughly 3\% of proposed mergers have faced extended review, and 32 proposals were dissolved or restructured (\cite{ConversableEconomist2025Antitrust}).

    Careful scrutiny of large-scale mergers is important. However, growing empirical and policy concern centers on unreported, small-scale mergers and non-merger affiliations. Research shows that small-scale mergers, often called “stealth mergers”, can reduce competition (\cite{majerovitz2021consolidation}; \cite{aggarwal2025stealth}; \cite{wollmann2019stealth}; \cite{wollmann2020get}; \cite{kepler2023stealth}). Non-merger affiliations, by contrast, encompass structural ties such as joint ventures, management consolidation, and overlapping boards of directors that do not trigger merger review. Among these, board interlocks are particularly noteworthy because they create direct governance links between firms, are explicitly addressed under the Clayton Act, and have drawn increasing policy attention. While board interlocks are documented (\cite{LemleyVanLoo2025OverlappingDirectors}; \cite{Poberejsky2025InterlocksInnovation}), their competitive effects remain largely hypothesized.

    This issue is especially relevant in health care, where consolidation is a persistent concern and several states have begun regulating small-scale mergers and non-merger affiliations (\cite{ncsl2024healthcare}). Additionally, in September 2025, the FTC announced the resignation of three directors from Sevita Health after uncovering board overlap with Beacon Specialized Living Services, both providers of residential care for individuals with intellectual and developmental disabilities, citing concerns under the Clayton Act and urging firms to proactively avoid such conflicts (\cite{ftc2025sevita}). Board interlocks are no longer abstract governance issues but active antitrust targets. Thus, in this paper, I construct a measure of non-merger affiliation among U.S. nonprofit hospitals that captures both interlocking boards and executive cross-board service, and I examine how these affiliations influence hospital behavior.

    I collect data on nonprofit hospital executives and board members from publicly available IRS Form 990 filings and merge these records with the American Hospital Association (AHA) Annual Survey. Using this combined dataset, I classify hospital affiliations based on system membership and geographic market. I define a non-merger affiliation as the presence of a common board member or an executive serving on another hospital’s board, where the two hospitals are otherwise unaffiliated. That is, they do not belong to the same system or network but share leadership ties. Because the hypothesized effects of non-merger affiliations are anti-competitive, whether the affiliated hospitals compete for the same patients is a critical distinction. To capture this, I separate affiliations into those occurring within the same Hospital Referral Region (HRR) and those across different HRRs, where HRRs represent geographic hospital markets (\cite{Wennberg1973SmallAV}). I show that [X]\% of nonprofit hospitals in my sample share a leader with an otherwise unaffiliated hospital in the same market, and [Y]\% of HRRs contain at least one such affiliation. I consider several hospital outcomes as behaviors that are informed by the corporate governance literature and hospital consolidation literature. I group these outcomes into categories of investment, patient composition, service offerings, and operating costs.

    Similarly to the merger literature, selection into the decision to become affiliated complicates causal inference. I begin with a staggered difference-in-differences design comparing hospitals that form an affiliation to those that do not, and I show that [INSERT FINDINGS]. While event-study estimates reveal no evidence of differential pre-trends, unobserved factors may still influence affiliation decisions. Therefore, my preferred specification employs a triple-differences approach that leverages variation across time, affiliation status, and whether the affiliation occurs within the same geographic market. This design isolates the incremental effect of affiliating with a hospital in the same market where competitive effects are most salient. Using this approach, I find that [INSERT FINDINGS].

    [IMPLICATIONS OF FINDINGS] Firms are responsive to antitrust regulation. That is, enforcement is a deterrant for anti-competitive behavior. Thus, it is important to think about whether regulation should increase even for small-scale affiliations. (something like this for policy implications?)

    Prior work establishes that corporations are deeply interconnected through overlapping directors and that these ties shape governance practices and firm performance (\cite{dooley1969interlocking}; \cite{ferris2003too}; \cite{field2013busy}; \cite{bizjak2009option}; \cite{bouwman2011corporate}; \cite{chiu2013board}; \cite{cai2014board}; \cite{baran2017director}; \cite{geng2021does}; \cite{schonlau2009board}; \cite{cai2012board}; \cite{cheng2021social}). That literature largely centers on publicly traded firms and frames overlap as either a governance friction (busy directors, option backdating, strategy convergence) or a source of private benefits (director networks that raise value, profitability, and post‑M\&A performance). In contrast, I study non‑merger affiliations in nonprofit hospitals, an arena often presumed less prone to anti-competitive behavior (\cite{baer2014clayton}; \cite{aai2013section7}) despite mounting evidence of competitive risks in health care (\cite{hulver2023ftc}). I broaden the construct beyond traditional board interlock to include executive cross‑board service, and I move the focus from governance proxies and firm‑level valuation to market‑relevant behaviors that are informed by the literature on both governance and consolidation (\cite{baran2017director}; \cite{geng2021does}; \cite{schonlau2009board}; \cite{cai2012board}; \cite{cheng2021social}). Methodologically, I leverage a triple‑differences design that distinguishes affiliations formed within versus across geographic markets to isolate the incremental, competition‑salient effect of overlap, rather than inferring behavior from financial outcomes alone.

    A large empirical literature examines the effects of consolidation in health care, particularly through hospital mergers and system integration. Early work found little evidence that consolidation improved operating costs, efficiency, or quality of care (\cite{alexander1996short}; \cite{ho2000hospital}; \cite{dranove2003hospital}). Yet from 2000 to 2020, hospital mergers surged, leaving a quarter of markets with no independent hospitals by 2020 (\cite{ElevanceHealth2023}). This trend spurred extensive research on price, quality, and efficiency. Studies consistently document modest cost and efficiency gains for hospitals joining larger systems (\cite{schmitt2017hospital}; \cite{andreyeva2024corporatization}; \cite{craig2021mergers}), but these gains rarely translate into lower prices; instead, mergers typically lead to significant price increases as hospitals gain bargaining power with insurers (\cite{gaynor2012impact}; \cite{boozary2019association}; \cite{cooper2019price}; \cite{andreyeva2024corporatization}). Evidence on quality is mixed: most studies find little effect on readmissions or mortality (\cite{haas2011mergers}; \cite{beaulieu2020changes}), though some report increased treatment intensity and mortality for heart disease patients (\cite{hayford2012impact}) or higher readmission rates overall (\cite{andreyeva2024corporatization}). Beyond price and quality, research highlights mechanisms such as convergence in treatment styles (\cite{eliason2020acquisitions}), consolidated service offerings (\cite{mariani2022impact}), and redistribution of patients within markets, including declines in Medicaid admissions in highly concentrated areas (\cite{desai2023hospital}). While this literature focuses on mergers and system affiliation, my study examines a different form of structural connection: non-merger affiliations through board and executive overlap.

    The paper proceeds as follows. In Section \ref{sec:background}, I provide institutional details on nonprofit hospital governance and hospital affiliations. In Section \ref{sec:data}, I outline the data collection process and provide summary statistics. In Section \ref{sec:simpledid}, I estimate a simple staggered difference-in-differences model leveraging affiliations over time. Finally, in Section \ref{sec:tripledid}, I leverage variation in both affiliation and geography to estimate a triple differences specification.

    \section{Background/Setting}\label{sec:background}

    \subsection{Nonprofit Hospital Governance}

    Nonprofits are governed at a high level by a board of directors: a group of unpaid community members that gather to oversee the broad direction and strategies of the firm. The board of directors does not mange the day-to-day activity of the hospital, but provides oversight to the highest level of manager, the Chief Executive Officer, whom the board also selects. The practical governance of boards is highly variable across different hospitals. State-level policies might influence details such as the regularity of board meetings or board member term limits. However, the main function of the board remains consistent: to provide strategic direction and oversight. In research focused on hospital governance, several factors have been found to be correlated with strategic change or hospital-level behaviors, including more influence over the CEO or top management (\cite{golden2001will}; \cite{alexander2008governance}; \cite{jiang2012enhancing}). Additionally, since the Affordable Care Act, boards have been actively involved in quality oversight (\cite{jha2010hospital}; \cite{prybil2014board}; \cite{prybil2010board}). Therefore, while boards are not usually involved in the day-to-day operations of hospitals, they do influence strategy and long-run hospital behavior.

    The board of directors selects a Chief Executive Officer, who then selects other members of the executive team such as a Chief Financial Officer and/or Chief Medical Officer. This leadership team manages the day-to-day operations of the hospital. An executive team usually consists of at least a Chief Executive Officer (CEO) and Chief Financial Officer (CFO), but there is variation in how firms organize these teams. Some hospital executives specialize in health care administration by earning a degree in health care management, or an MBA specific to health care.

    In this paper, I consider the effects of board and executive interlock, someone sitting as an executive or on the board of multiple organizations. Board overlap does occur in health care with instances of organizations denied merger permission due to boards not acting independently (\cite{huberfeld2006tackling}), of hospitals declaring informal affiliations through leadership (\cite{barnett_babcock_2012}), and of life sciences companies with substantial documented overlap (\cite{manjunath2024illegal}). For many years, oversight of this activity in nonprofits was considered second-order to that of for-profits. However, in 2024, the FTC raised concern over board interlock of two prominent health care organizations (\cite{ftc2025sevita}). Additionally, board will sometimes ask executives from other organizations to sit on their board (\cite{gordon2025ftc}; \cite{alston2024governance}).

    Early research on board interlock highlights that firms are not isolated entities but interconnected through shared directors (\cite{dooley1969interlocking}), with interlocks linked to firm size, financial ties, and management control. Research has examined governance implications, including whether overlapping boards overextend directors (\cite{ferris2003too}; \cite{field2013busy}), increase stock option backdating (\cite{bizjak2009option}), or promote convergence in governance strategies (\cite{bouwman2011corporate}; \cite{chiu2013board}; \cite{cai2014board}). Empirical evidence focuses largely on financial outcomes: well-connected directors can enhance firm value (\cite{baran2017director}), and board overlap is associated with higher profitability (\cite{geng2021does}), greater acquisition activity, and improved post-merger performance (\cite{schonlau2009board}; \cite{cai2012board}). Mechanisms include shared IT investments (\cite{cheng2021social}), reduced R\&D, and increased product differentiation (\cite{geng2021does}).

    \subsection{Hospital Affiliations}\label{sec:hospaffil}

    Board interlock is often happening in conjunction with affiliation agreements: any relationship between otherwise independent health care providers designed to create shared advantage and value (\cite{aha2018affiliations}). Some examples of affiliation agreements include joint ventures for service provision, clinical agreements among hired health care workers, or joint management. It is common for these types of agreements to involve some sort of shared governance (\cite{aha2018affiliations}; \cite{natlawreview2023ftc}). While some have raised anti-trust concerns over affiliation agreements, they are largely undocumented, and therefore their effects are unknown (\cite{rand2023consolidation}). Organizations do not require pre-authorization to form affiliation agreements, but there are several states that now require notice of any type of affiliation (\cite{ncsl2024healthcare}).

    The more documented and studied type of hospital affiliation is that of traditional mergers and acquisitions. I will discuss the literature surrounding this topic in order to inform relevant behaviors that could also be affected by affiliation agreements. Hospital consolidation has accelerated over the past two decades, with mergers becoming so prevalent that by 2020, a quarter of hospital markets had no independent hospitals (\cite{ElevanceHealth2023}). Early studies found little evidence that consolidation improved operating costs, efficiency, or quality (\cite{alexander1996short}; \cite{ho2000hospital}; \cite{dranove2003hospital}), but more recent work documents modest efficiency gains for hospitals joining larger systems (\cite{schmitt2017hospital}; \cite{andreyeva2024corporatization}; \cite{craig2021mergers}). These gains, however, rarely translate into lower prices; instead, mergers consistently lead to significant price increases due to enhanced bargaining power with insurers (\cite{gaynor2012impact}; \cite{boozary2019association}; \cite{cooper2019price}; \cite{andreyeva2024corporatization}). Evidence on quality is mixed: most studies find little change in readmission or mortality rates (\cite{haas2011mergers}; \cite{beaulieu2020changes}), though some report higher treatment intensity and mortality for heart disease patients (\cite{hayford2012impact}) and increased readmissions overall (\cite{andreyeva2024corporatization}). Other research highlights mechanisms such as convergence in treatment styles (\cite{eliason2020acquisitions}), changes in service offerings (\cite{mariani2022impact}), and shifts in admission patterns, including reduced Medicaid admissions in concentrated markets (\cite{desai2023hospital}). These findings underscore that consolidation can influence not only costs and prices but also care delivery and patient access.

    \section{Data and Summary Statistics}\label{sec:data}

   To create a measure of board/executive overlap, I compile data on nonprofit hospital board of directors and executives in years 2013-2023 from publicly available Tax Form 990s. All sufficiently large nonprofits must file these forms with the Internal Revenue Service (IRS) each year. They contain a section in which nonprofits declare their board, executives, and highest compensated employees. This paper focuses only on board members and executives. I limit to firms that fill out Schedule H of the tax form, an indicator that the firm operates a hospital. I extract all names and positions of individuals listed in the section titled ``Officers, Directors, Trustees, Key Employees, and Five Highest Compensated Employees". I also record information on the hospital entity: the Employee Identification Number (EIN), organization name and location, and other characteristics. The initial extraction yields 2,096 nonprofit hospitals.

   Next, I match hospitals in the tax form data to hospitals in other publicly available data sources. The only common information about hospitals across these sources is the name and location. Therefore, I use fuzzy string matching methods to create a crosswalk of EINs to identification numbers in the American Hospital Association (AHA) survey data. Specifically, I compute the Jaro-Winkler distance between hospital names in each data set. Then, I record matches with sufficiently similar names and addresses, yielding a sample of 1,824 matched hospitals.

    I maintain information about the individuals in the data, including whether they have a title for medical doctor, nurse, or having a degree in health administration. To clean the names, I remove any titles or credentials, and I convert common nicknames to their longer versions for consistency. I also combine names within the same firm that have either slight misspelling differences or, if middle names are only sometimes listed, they have at least 2 names in common. For example, if John Lee Matthew is a board member of hospital A in 2017, and John Matthew is a board member of hospital A in 2018, I assume these names represent the same person. Finally, I use the gender package in R to predict each individual's gender using historical data (\cite{gender}).

    I present summary statistics of characteristics of the people in the data in Table \ref{tab:board_people}. There are just under 52,000 people in the data at some point over the course of 10 years, 63\% of which are board members and the remaining 37\% are executives. Board members are slightly more likely to be medical doctors than executives, but the fraction of doctor board members is still relatively low at 18\%. Nurses are consistent across board and executives, but only 3\% of the sample are nurses. A negligible number of people claim to have a health administration degree. Executives are slightly more likely to be female according to the name algorithms, with 38\% of executives estimated to be female and 30\% of board members estimated to be female. I also include average team composition of hospital boards and executive teams, where these trends largely hold at the hospital level. On average, hospital boards contain 19 members and executive teams contain 14 members.

    \import{Objects}{boardandexec_people.tex}

    I create a list of overlapping board member or executive candidates by identifying any names associated with multiple hospital in the same year. However, it is very common for hospitals within a system to largely share the same board members and/or executives. This is not the type of variation I seek to investigate in this paper, since hospitals within systems are already heavily affiliated. Therefore, I define the relevant type of overlap as someone being on the leadership team of multiple hospitals, where the hospitals are otherwise unaffiliated by system or network. Hospitals within systems are still eligible to have overlap with other hospitals that do not belong to the same system. I exclude hospitals that gain or lose overlapping board members several times in the sample period.

    In the hospital merger literature, mergers have been found to be more significant when the two hospitals involved are located geographically close (\cite{cooper2019price}). Therefore, I define two types of overlap: overlap with an otherwise non-affiliated hospital in the same market, or overlap with an otherwise unaffiliated hospital in a different market. I use the market definitions developed by Dartmouth Atlas, Hospital Referral Regions (HRRs), which are meant to capture geographic regions of the average patient's choice set of hospitals. From these overlap definitions, I characterize hospitals by the first and last year they are connected in the same HRR, and the first and last year they are connected in a different HRR.

    \subsection{Documenting Leadership Overlap}

    In Figure \ref{fig:connected_percent}, I illustrate the prevalence of overlapping board members or executives over time. The fraction of hospitals with overlapping leaders rises gradually from 10\% in 2014 to 15\% in 2017, then declines slightly after 2017. The peak in 2017 lacks an obvious explanation, but the subsequent decline may reflect growing consolidation in health care, which reduces opportunities to connect with hospitals outside the same system. I also plot the fraction of markets that include at least two hospitals with overlapping leaders. This measure increases from 15\% in 2014 to nearly 30\% in 2017, then falls to roughly 20\% by 2023. For comparison, about 60\% of hospitals in my sample begin as part of a system, and this fraction rises steadily over time. Although informal affiliations through overlapping leaders occur less often than formal system membership, they remain a notable feature of the U.S. hospital landscape. I also present the geographic distribution of these pairs in Figure \ref{fig:connected_maps}. The blue lines represent hospitals that share a common board member or executive, and the red dots are all other hospitals. Connected hospitals are distributed fairly evenly across the US, apart from relatively few connections on the west coast relative to the population.

    \begin{figure}[ht!]
        \centering
        \vspace{9mm}
        \caption{Prevalence of Board/Executive Interlock}
        \includegraphics[width=.9\textwidth]{Objects/connected_percent.pdf}
        \label{fig:connected_percent}
    \end{figure}

    \begin{figure}[ht!]
        \centering
        \caption{Geographic Distribution of Board/Executive Interlock}
        \includegraphics[width=.85\textwidth]{Objects/connected_maps.pdf}
        \label{fig:connected_maps}
    \end{figure}

    \subsection{Characteristics of Hospitals with Overlap}

    Hospitals in the sample fall into three categories: always connected, never connected, or connected at some point in time. Table \ref{tab:hosp_group_stats} reports characteristics for each group. Hospitals that never exhibit leadership overlap, whether within or across markets, differ markedly from connected hospitals, with substantially fewer beds and nurses. In contrast, hospitals that become connected during the sample period resemble connected hospitals much more closely. Over 300 hospitals gain overlap within the same HRR, and almost 400 gain overlap across different HRRs. These groups are not mutually exclusive, so some hospitals experience overlap in both settings.\footnote{I account for this in the estimation.} Nearly 600 hospitals remain connected across different HRRs throughout the sample, making this form of overlap the most common.

    \import{Objects}{hosp_group_stats.tex}

    I now present characteristics of connected pairs of hospitals in Table \ref{hospital_pair_types}. In the first column, I show pair characteristics of hospitals connected within the same HRR. In the second column, I present characteristics of hospitals connected to otherwise unaffiliated hospitals located in a different HRR. Therefore, all hospital pairs in the table are connected, but only the first column are connected to hospitals that are potentially direct competitors. Each hospital in a pair is general or specialty (I only consider adult hospitals for these categories), children or adult, and independent or belonging to a system, all characteristics in the AHA data. Additionally, the median bed size of all hospitals in the sample is 119, and I define hospitals as being either small or large depending on whether they fall below or above this median. There are typically 1.7 connections in the same HRR and almost 8 connections in different HRRs.

    \import{Objects}{hospital_pair_types.tex}

    I will first discuss hospital pairs located in the same HRR. Most of these pairs, 64\%, consist of two general hospitals, and are thus competitors in terms of typical services provided. An additional 12\% of these pairs are between a general and a specialty hospital, and the remaining 1\% are between two specialty hospitals. Similarly, three quarters of the pairs in the same HRR are between two adult hospitals. While the largest fraction of these pairs, 44\%, is between two independent hospitals, there is also a significant portion, 40\%, that occur between one independent hospital and one system-affiliated hospital. Only 16\% of these pairs are between two hospitals belonging to different systems. Almost half of the pairs are between one small and one large hospitals, and the average bed difference is over 200, indicating a significant size difference on average. Finally, pairs within the same HRR are typically 35 miles apart.

    Now, I discuss hospitals that share leaders with hospitals in different markets. 80\% of these pairs are between two general hospitals, and 90\% are between two adult hospitals. Thus, it is likely that hospitals are seeking expertise from other hospitals similar to theirs moreso than in the same HRR. There is a fairly even spread of whether the hospitals in the pair are independent. Still, 50\% of pairs are between one large and one small hospital, and the bed size difference is 238 on average. As expected, the hospitals in these pairs are located much farther apart geographically than those in the same HRR.

    Hospital decisions to become connected are obviously endogenous. Thus, when thinking about how overlap affects hospital behavior, the relevant comparison group is important. To think about this further, I present trends over time for a select number of outcomes for four different groups of hospitals: (1) those that gain overlap in the same HRR, (2) those that gain overlap in a different HRR, (3) those that are always connected in a different HRR, and (4) those that are never connected. I show these trends for outcomes of number of discharges, expenses, and number of services offered by the hospital in Figure \ref{fig:outcome_desc_graph}.

    \begin{figure}[ht!]
        \centering
        \caption{Select Outcomes by Overlap Status}
        \includegraphics[width=.7\textwidth]{Objects/outcome_desc_graph.pdf}
        \label{fig:outcome_desc_graph}
    \end{figure}

  These trends show that not all hospitals serve as a suitable comparison group for those that gain a connection within the same HRR. In the top panel, which plots the average number of services over time, hospitals that remain unconnected or always connected in a different HRR follow a different trajectory than hospitals that become connected. A similar pattern appears for discharges: hospitals that gain overlap track a comparable path before the connection, then diverge, while hospitals that never gain overlap exhibit a distinct trend of declining Medicare and Medicaid discharges. Finally, total operating expenses reveal entirely different slopes for hospitals that never gain overlap compared to those that do, whether in the same or a different market. For this reason, when estimating the effect of gaining overlap in the same market, I compare these hospitals only to those that become connected in a different market. This approach allows me to isolate the effect of connecting with a competitor from the broader effect of becoming connected.

    \section{Estimation and Identification}

    I estimate the effect of gaining or losing leadership overlap on several hospital outcomes. Since hospitals become connected via leadership at different times and their effects could be heterogeneous, I employ a stacked difference-in-differences (DiD) approach. Specifically, I construct event-time observations for each treated hospital around the time of gaining or losing overlap and stacks these into a single panel. This method accommodates variation in treatment timing across hospitals and enables estimation of dynamic effects. On the stacked panel, I estimate the following equation:

    \begin{equation}\label{eq:stacked}
    Y_{it} = \alpha + \sum_{k \neq -1} \beta_k \cdot D_{i,t+k} + \gamma_i + \delta_t + \varepsilon_{it}
    \end{equation}

    \noindent where $Y_{it}$ is the outcome for hospital $i$ at time $t$ (closure or merger indicator), $D_{i,t+k}$ is an indicator for event time $k$ relative to the change in leadership overlap (with $k = -1$ as the omitted category), $\gamma_i$ and $\delta_t$ are hospital and time fixed effects, respectively, and $\beta_k$ captures the effect at each event time $k$.

    \section{Leadership Overlap and Consolidation}

    While hospitals often publicly declare that their leadership affiliations are not intended to lead to mergers or other consolidation (add citations), the underlying nature of overlapping leadership with regards to other types of hospital affiliations is an open empirical question. Theoretically, there are three ways that leadership overlap and consolidation could be related. First, leadership overlap may act as a substitute for formal mergers. In this case, hospitals leverage shared governance to achieve some benefits of consolidation, such as strategic alignment, resource coordination, and cost efficiencies, without incurring the regulatory scrutiny and transaction costs associated with a merger. If this hypothesis holds, we would expect leadership overlap to reduce the likelihood of subsequent merger activity. Second, leadership overlap may serve as a precursor to merging. Hospitals might begin by sharing board members or executives to build trust, harmonize decision-making, and prepare for operational integration. Under this interpretation, overlap is an intermediate step toward full consolidation, and its presence should increase the probability of a future merger. Third, leadership overlap may result from intentions to merge. Here, overlap is not an independent strategic choice but rather a signal of an already-decided merger plan. This scenario introduces endogeneity concerns, as leadership overlap and merger intentions are jointly determined. In addition to its possible role in mergers, leadership overlap may also influence the likelihood of hospital closure. Shared leadership could provide access to broader networks, financial resources, and managerial expertise, which might help struggling hospitals remain viable and avoid closure. Conversely, overlap could signal strategic realignment within a system, where weaker hospitals are deliberately phased out to concentrate resources in stronger facilities. In this sense, leadership overlap could either reduce closure risk by stabilizing operations or increase it if consolidation strategies involve selective shutdowns.

    The estimates from Equation \ref{eq:stacked} where hospital closure is the outcome variable are shown in Figure \ref{fig:closure}. I consider three types of variation in leadership overlap: gaining overlap with a hospital in the same HRR or different HRR, and losing overlap with a hospital in the same HRR or different HRR. While gaining leadership overlap with another hospital in the same or different HRR is not correlated with closing in the future, hospitals that lose overlap are .15 percentage points more likely to close 1-4 years after losing connectedness compared to (who are the comparison hospitals here??). Part of this may is mechanical, as hospitals that close (drop out of the data) necessarily lose board overlap. (I don't think theres any way to get rid of this mechanical relationship...)

    \begin{figure}[ht!]
        \centering
        \vspace{9mm}
        \caption{Effect of Overlap on Closure}
        \includegraphics[width=.9\textwidth]{Objects/closure.pdf}
        \label{fig:closure}
    \end{figure}

    I also estimate Equation \ref{eq:stacked} where the outcome variable is an indicator for whether or not the hospital is independent, where independent means the hospital does not belong to a system. I present the results of this estimation in Figure \ref{fig:independent}. In this case, losing leadership overlap is not associated with any changes in system status. However, gaining leadership overlap in the same HRR is slightly negatively associated with being independent in the first year after becoming connected via leadership.

    \begin{figure}[ht!]
        \centering
        \vspace{9mm}
        \caption{Effect of Overlap on Being Independent}
        \includegraphics[width=.9\textwidth]{Objects/independent.pdf}
        \label{fig:independent}
    \end{figure}

    This does not necessarily mean that leadership overlap causes closure or mergers, as the relationship could be more complex than that. Add more to this paragraph about how this is not causal but indicates that the nature of board overlap is more than just expertise sharing??

    \section{Leadership Overlap and Hospital Outcomes}

    If leadership overlap, encompassing informal hospital affiliations, is a substitute for more formal mergers, there are several outcomes that we typically see affected in the hospital merger literature that we can test whether they are also associated with leadership overlap. These outcomes fall into categories of patient population, service offerings, and operating expenses.

    \subsection{Patient Population}

    \begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.47\textwidth}
        \caption{Medicaid Discharges}
        \includegraphics[width=\textwidth]{Objects/mcaid_discharges.pdf}
        \label{fig:mcaid}
    \end{subfigure}
    \begin{subfigure}[b]{0.47\textwidth}
        \caption{Medicare Discharges}
        \includegraphics[width=\textwidth]{Objects/mcare_discharges.pdf}
        \label{fig:mcare}
    \end{subfigure}
    \begin{subfigure}[b]{0.47\textwidth}
        \caption{Total Discharges}
        \includegraphics[width=\textwidth]{Objects/tot_discharges.pdf}
    \end{subfigure}
    \label{fig:mcaid_mcare}
    \end{figure}

    \subsection{Service Offerings}

    \begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.47\textwidth}
        \caption{Bed Concentration}
        \includegraphics[width=\textwidth]{Objects/bed_conc.pdf}
        \label{fig:bed_conc}
    \end{subfigure}
    \begin{subfigure}[b]{0.47\textwidth}
        \caption{Number of Services}
        \includegraphics[width=\textwidth]{Objects/num_services.pdf}
        \label{fig:num_services}
    \end{subfigure}
    \label{fig:services}
    \end{figure}

    \subsection{Investment and Operating Expenses}

    \begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.47\textwidth}
        \caption{Purchases (Building or Equipment)}
        \includegraphics[width=\textwidth]{Objects/any_purch.pdf}
        \label{fig:any_purch}
    \end{subfigure}
    \begin{subfigure}[b]{0.47\textwidth}
        \caption{Total Operating Expenses}
        \includegraphics[width=\textwidth]{Objects/tot_operating_exp.pdf}
        \label{fig:tot_oper_exp}
    \end{subfigure}
    \label{fig:expenses}
    \end{figure}

    \section{Conclusion}

    This study contributes to the growing literature on hospital governance and market structure by examining the extent and implications of board member overlap across hospitals. By identifying the prevalence of shared board membership, I provide new insights into how governance ties may influence competition, decision-making, and resource allocation in the hospital sector. This paper sheds light on the role of board interlocks in potentially shaping hospital behavior, offering an important perspective on the non-ownership mechanisms that can link hospitals within the same market.

    Using a comprehensive dataset of board members in nonprofit hospitals from 2017-2022, I document the characteristics of hospitals with shared board members and analyze how these overlaps have evolved over time. I identify connections between hospitals and examine the patterns of overlap by hospital type, market, and geographic region. The results indicate that a significant share of hospitals are connected through board interlocks, and a significant portion of hospital markets contain board overlap within the market. Further, I show geographic distributions of overlap and characterize connected hospital pairs by important characteristics such as ownership, size, and specialty.

    These findings hold important implications for policymakers and future research. Regulators and antitrust authorities should consider the potential competitive effects of board member overlap when assessing hospital market dynamics, particularly in regions with high consolidation or limited provider choice. Future research could further investigate how these governance connections influence hospital pricing, quality of care, and patient outcomes. Expanding this work to explore causal relationships and the broader economic consequences of board interlocks would provide valuable insights for competition in health care.

    \newpage

    \printbibliography

    \appendix

\begin{figure}
    \centering
    \caption{Outcome Variables Over Time, Patient and Service Allocation}
    \includegraphics[width=0.75\linewidth]{Objects/outcome_perc_dis_bed_graph.pdf}
    \label{fig:outcome_means1}
\end{figure}

\begin{figure}
    \centering
    \caption{Outcome Variables Over Time, Purhcases, Costs, and Revenue}
    \includegraphics[width=0.75\linewidth]{Objects/outcome_purch_oper_rev_graph.pdf}
    \label{fig:outcome_means2}
\end{figure}

\end{document}